SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-04-10 2017-Xudong-SMT/ActaInformatica.tex
Refine proof 2&3

 Let $pn=$ be a pNet. The set of states 
 managed by the two algorithms both are subsets of the cartesian
 product of the PLTS states, as defined in page \pageref{def-states}. In a
 similar way, the set of state variables attached to each state are
 also equal. Comparing  
 transitions requires a little more care, because within the
 $MakeTransitions$ function, fresh variables are created everytime a
 synchronisation vector is used (this occurs within the $MkPred$ and
 $matching$ functions). However, given a global state, the call of
 $MakeTransitions$ will always return exactly the same set of OTs, up to a
renaming of these fresh variables. When this is not ambiguous, we will
speak of equality (of transitions, and of open automata) without the phrase ``up to a renaming of fresh variables''.

@@

  By induction hypothesis, we get $S_{n} \in States(L1_{reach})$, so during the
  call of $makeReachableSubAutomaton$ in Algorithm~\ref{alg1}, line
  12, there exists an OT
  $S_n \xrightarrow{l_{n}}S$ in $L1_{sat}$, with $l'_n=Rename(l_n)$;
  this OT will be checked and found SAT, so
  $S \in States(L1_{reach})$.

  Conversely, if $|S|_{L1_{reach}}=n+1$, we get a similar chain in
  $L1_{reach}$, with the transition $S_n \xrightarrow{l_{n}}S$.  Then $ S_n \in
  States(L1_{reach})$ and $l_n$ is satisfiable. By induction, we have
  $S_n \in States(L2_{reach})=States(L2_{sat})$, so there exists
  $l'_n=Rename(L_n)$ that will have
  been checked on the fly, in line 6 of Algorithm~\ref{alg3}. Then
  the transition $S_n \xrightarrow{l'_{n}}S$ will be added to
  $L2_{sat}$, and thus $S \in States(L2_{reach})$.

@@

There are two reasons why an OT would disappear during one application
of $Aut^k = Refine(Aut^{k-1})$: either its predicate, checked with the set of
$Assign(L^{k-1})$, became Unsat; or this OT is at
the end of a sequence starting from a state in $s_0\ \in State(L^k)$,
the next OT became Unsat, and the following states in the sequence are
not reachable using satisfiable OTs in $Aut^k$. More formaly, there is a chain
\[s_0\xrightarrow{ot_1}s_1\xrightarrow{ot_2}...s_{n}\xrightarrow{ot_{n+1}}s_{n+1}\]
with $Pred(ot_1,L^{k-1})$ Unsat, and all states $s_i, i \in [1..n]$
unreachable in $Aut^k$.

@@

\medskip
Now we prove that the triples at the end of the fixpoint computation
in both algorithms are equal.
Let $Aut^*_1$ and $Aut^*_2$ be two fixed points with $Refine(Aut^*_1) =
Aut^*_1$ and $Refine(Aut^*_2) = Aut^*_2$.  
We want to prove $L_1^* = L_2^*$.

We already have $Aut_2^0\ \subseteq Aut_1^0$.
We will prove in 2 steps:
first prove by induction that $Hyp^k = (L_2^k\ \subseteq L_1^k)$ is
preserved by application of $Refine$ (meaning that the refine function
in algorithm 1 cannot go ``faster'' than in algorithm 3), then prove
that when the fix-points are reached, then they are the same.

Suppose $Hyp^{k-1}$, and prove $Hyp^k$.
Consider an OT $l\ \in L_2^k$, but $l\ \notin L_1^k$.
We have \[ l\ \in L_2^k \implies l\ \in L_2^{k-1} \implies l\ \in
L_1^{k-1}\] so the transtion $l$ has been removed from $L_1^{k-1}$ during the last
iteration $Refine_1^k$. This can be caused by:
\begin{itemize}
\item
  either $Pred(l,L_1^{k-1})$ is Unsat, because of some assignment
    $u \in Assigns(L_1^{k-1})$. But $l \in L_2^k$, so its predicate
    $Pred(l,L_2^{k-1})$ is satisfiable; this is a contradiction,
    because $Assigns(L_2^k) \subseteq Assigns(L_1^k)$ (having
    more possible assignments cannot reduce satisfiability).
  \item
    or there is a sequence
    $s_0\xrightarrow{l_1}s_1\xrightarrow{l_2}...s_{n}\xrightarrow{l}s_{n+1}$,
    with $s_0 \in States(L_1^{k-1})$, $Pred(l_1,L_1^{k-1})$ is Unsat, and
    all further states unreachable in $L_1^{k-1}$. But as in the
    previous case, we would have $l_1$ in $L_2^k$ and not in $L_1^k$,
    reaching a similar contradiction.
\end{itemize}

\smallskip
Now consider the fixpoints of the $Refine$ iterations. From the
initial inclusion $L_2^0\ \subseteq L_1^0$, and our invariant, we get
inclusion of the fixpoints $L_2^*\ \subseteq L_1^*$.

Suppose there exists an OT $l$ in $L_1^*$ that is not in $L_2^*$. This
means that $Pred(l,Assigns(L_1^*))$ is satisfiable, but
$Pred(l,L_2^*)$ is satisfiable, that is a contradiction; or as in the
invariant proof, there is a similar sequence, with an OT $l_1$
bringing the contradiction.

>>>>

 Let $pn$ be a pNet. The set of states 
 managed by the two algorithms both are subsets of the cartesian
 product of the pLTS states, as defined in page \pageref{def-states}. In a
 similar way, the set of state variables attached to each state are
 also equal. Comparing  
 transitions requires a little more care, because within the
 $MakeTransitions$ function, fresh variables are created everytime a
 synchronisation vector is used (this occurs within the $MkPred$ and
 $matching$ functions). However, given a global state, the call of
 $MakeTransitions$ will always return exactly the same set of OTs, up to a
renaming of these fresh variables. When this is not ambiguous, we will
speak of equality (of transitions, and of open automata) without the phrase ``up to a renaming of fresh variables''.

@@

  By induction hypothesis, we get $S_{n} \in States(L1_{reach})$, so during the
  call of $makeReachableSubAutomaton$ in Algorithm~\ref{alg1}, line
  12, there exists an OT
  $S_n \xrightarrow{l_{n+1}}S$ in $L1_{sat}$, with $l'_{n+1}=Rename(l_{n+1})$;
  this OT will be checked and found SAT, so
  $S \in States(L1_{reach})$.

  Conversely, if $|S|_{L1_{reach}}=n+1$, we get a similar chain in
  $L1_{reach}$, with the transition $S_n \xrightarrow{l_{n+1}}S$.  Then $ S_n \in
  States(L1_{reach})$ and $l_{n+1}$ is satisfiable. By induction, we have
  $S_n \in States(L2_{reach})=States(L2_{sat})$, so there exists
  $l'_{n+1}=Rename(l_{n+1})$ that will have
  been checked on the fly, in line 6 of Algorithm~\ref{alg3}. Then
  the transition $S_n \xrightarrow{l'_{n+1}}S$ will be added to
  $L2_{sat}$, and thus $S \in States(L2_{reach})$.

@@

There are two reasons why an OT would disappear during one application
of $Aut^k = Refine(Aut^{k-1})$: either its predicate, checked with the set of
$Assign(L^{k-1})$, became UNSAT; or this OT is at
the end of a sequence starting from a state in $s_0\ \in State(L^k)$,
the next OT became UNSAT, and the following states in the sequence are
not reachable using satisfiable OTs in $Aut^k$. More formaly, there is a chain
\[s_0\xrightarrow{ot_1}s_1\xrightarrow{ot_2}...s_{n}\xrightarrow{ot_{n+1}}s_{n+1}\]
with $Pred(ot_1,L^{k-1})$ UNSAT, and all states $s_i, i \in [1..n]$
unreachable in $Aut^k$.

@@

\medskip
Now we prove that the triples at the end of the fixed point computation
in both algorithms are equal.
Let $Aut^*_1$ and $Aut^*_2$ be two fixed points with $Refine(Aut^*_1) =
Aut^*_1$ and $Refine(Aut^*_2) = Aut^*_2$.  
We want to prove $L_1^* = L_2^*$.

We already have $Aut_2^0\ \subseteq Aut_1^0$.
We will prove in 2 steps:
first prove by induction that $Hyp^k = (L_2^k\ \subseteq L_1^k)$ is
preserved by application of $Refine$ (meaning that the refine function
in algorithm 1 cannot go ``faster'' than in algorithm 3), then prove
that when the fixed points are reached, then they are the same.

Suppose $Hyp^{k-1}$, and prove $Hyp^k$.
Consider an OT $l\ \in L_2^k$, but $l\ \notin L_1^k$.
We have \[ l\ \in L_2^k \implies l\ \in L_2^{k-1} \implies l\ \in
L_1^{k-1}\] so the transtion $l$ has been removed from $L_1^{k-1}$ during the last
iteration $Refine_1^k$. This can be caused by:
\begin{itemize}
\item
  either $Pred(l,L_1^{k-1})$ is UNSAT, because of some assignment
    $u \in Assigns(L_1^{k-1})$. But $l \in L_2^k$, so its predicate
    $Pred(l,L_2^{k-1})$ is satisfiable; this is a contradiction,
    because $Assigns(L_2^k) \subseteq Assigns(L_1^k)$ (having
    more possible assignments cannot reduce satisfiability).
  \item
    or there is a sequence
    $s_0\xrightarrow{l_1}s_1\xrightarrow{l_2}...s_{n}\xrightarrow{l}s_{n+1}$,
    with $s_0 \in States(L_1^{k-1})$, $Pred(l_1,L_1^{k-1})$ is UNSAT, and
    all further states unreachable in $L_1^{k-1}$. But as in the
    previous case, we would have $l_1$ in $L_2^k$ and not in $L_1^k$,
    reaching a similar contradiction.
\end{itemize}

\smallskip
Now consider the fixed points of the $Refine$ iterations. From the
initial inclusion $L_2^0\ \subseteq L_1^0$, and our invariant, we get
inclusion of the fixed points $L_2^*\ \subseteq L_1^*$.

Suppose there exists an OT $l$ in $L_1^*$ that is not in $L_2^*$. This
means that $Pred(l,L_1^*)$ is satisfiable, but
$Pred(l,L_2^*)$ is not satisfiable, that is a contradiction; or as in the
invariant proof, there is a similar sequence, with an OT $l_1$
bringing the contradiction.
